Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-05-09 VegOzone.tex
First revision of sec 3.2.2

\subsubsection{Ozone}
\label{subsubsec:anomal_ozone}
As we discussed in the preceding Section~\ref{subsubsec:anomal_tpq}, meteorological conditions in 2018 were significantly different from the climatological average. How does this translate into observed ozone anomalies \chem{\Delta[O_3]}?

If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the recent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance. In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long-term observations at Esrange, Jergul/Karasjok, and Pallas. For Svanvik this assessment is based on rather uncertain observations from 1986--1996. We also discussed the general patterns, similarities, and differences. We shall now investigate \chem{\Delta[O_3]} in 2018/19 for these climatologies in more detail.

In the following, we will use the interchangeable terms anomalies and residuals to distinguish between a climatological (anomalies) and a statistical (residuals) point of view.
All the above factors make it necessary to look not only at anomalies but statistically also to assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison. Since the climatology of the latter is more reliable, we will start with ground-level ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018. As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.

In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown. Both stations display a clear, but not significant, deviation from the climatological mean. On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively. Both distributions are slightly skewed towards positive deviations with $\frac{1}{4}$ of days (3. quantile, q3) falling above $5.8\,\unit{ppb}$ and $5.6\,\unit{ppb}$, respectively.

Supplement Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandia climatology. For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively. While the probability density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
With respect to the climatology for Svanvik (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means, but daily average ozone concentrations are slightly elevated in 2018 compared to 2019.

The \chem{\left<\Delta\left<[O_3]\right>\right>} for Svanvik in 2018 is slightly larger than for Esrange and Pallas. This could mean either that 2018 had been more exceptional at Svanvik or the derived climatology may need a correction for increased tropospheric background \chem{[O_3]}. Based on the meteorological conditions, we assume 2019 as an average year which leads to an estimated constant bias of about $1.2\,\unit{ppb}$ in \chem{[O_3]}. Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$ \citep{NILU2003, BER:Ruoho-Airola2015}. Former is compatible with our assumed bias for the climatology derived for Svanvik.

To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question. Results for Esrange and Pallas are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}. This t-test clearly shows, that at both sites the residuals are significant on the $1\sigma$ level for at least $50\,\unit{\%}$. In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed. From Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandia climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019. In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 for the climatology from Svanvik. We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.

In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and estimate the ozone concentration anomalies for 2018/19. In Fig.~\ref{fig:anomalies_svanvik}d), the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month. Negative deviations from climatology are shown as a negative percentage. Note that missing bars for certain months are both due to data availability (see Fig~\ref{fig:data_svanvik_2018_2019} for details concerning Svanvik) and anomalies not meeting the criteria of $\pm 1\sigma$ deviation from the climatological mean. In April/May 2018, ozone was significantly enhanced, while 2019 was generally an average year. In both years, ozone concentrations were significantly below average in November. July 2018 represented by the color bar has not been corrected for the missing data due to data acquisition problems. The corrected value is indicated by the star. A detailed describtion of the gap-filling based on a Reynolds's decomposition method will be published seperately.

To conclude, ozone concentrations in 2018 have been elevated compared to the respective climatologies, however, on average not significantly by statistical means. At Esrange and Pallas at least $25\,\unit{\%}$ (3.~quantile) of anomalies are significant on a $3\sigma$ level or more, while at Svanvik the 3.~quantile lies at about $2\sigma$. This means that in 2018 the stations at Esrange and Pallas saw more days with elevated ozone than was monitored at Svanvik.

>>>>

\subsubsection{Ozone}
\label{subsubsec:anomal_ozone}
To assess the significance of 2018/19 anomalies, we estimate the bias due to the change in the tropospheric background \chem{[O_3]} since the 1990s and account for the missing data in June 2018. To reconstruct these data, we performed a Reynold's decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone obserations at Esrange and Pallas (compagnion paper). In the following, we will describe the bias-correction.

In 2018, the probability density function of ozone concentration anomalies (\chem{\Delta[O_3]}) for Esrange and Pallas deviate clearly, but not significantly from the climatological mean (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}). On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively. Both distributions are slightly skewed towards positive deviations with $\frac{1}{4}$ of days (3. quantile, q3) falling above $5.8\,\unit{ppb}$ and $5.6\,\unit{ppb}$, respectively.

Similarly, we find a skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}). In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and the annual average anomally in 2019 can be used for bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>} = $(1.2\pm 5.0)\,\unit{ppb}$). To further strengthen this, we look at trends in ozone observations at Esrange and Pallas and compute the average difference between recent years and the mid 1990s (\chem{\Delta_{present-past}\left<[O_3]\right>}). Present is defined as the average over 2010--2018 and past as the time between the respective beginning of data taking until 1997. We do not find a clear trend. For Esrange \chem{\Delta_{present-past}\left<[O_3]\right>} amounts to $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas to $(-2.5\pm 3.7)\,\unit{ppb}$ \citep[][for comparison]{NILU2003, BER:Ruoho-Airola2015}. The observed trend at Esrange is compatible with our assumed bias correction for the Svanvik ozone climatology.

%To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question. Results for Esrange and Pallas are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}. This t-test clearly shows, that at both sites the residuals are significant on the $1\sigma$ level for at least $50\,\unit{\%}$. In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed. From Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandia climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019. In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 for the climatology from Svanvik. We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.

In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and assess the monthly significance of the ozone concentration anomalies in 2018/19 (Fig.~\ref{fig:anomalies_svanvik}d)). The resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month in Fig~\ref{fig:data_svanvik_2018_2019}. Negative deviations from climatology are shown as a negative percentage. Note that missing bars for certain months are both due to data availability and anomalies not meeting the criteria for significance. The the star indicates resulting anomalies in after the reconstruction of missing values in July. In April/May 2018, ozone was significantly enhanced, while no significant enhancement is found in 2019. In both years, ozone concentrations were significantly lower than average in November.

%To conclude, ozone concentrations in 2018 have been elevated compared to the respective climatologies, however, on average not significantly by statistical means. At Esrange and Pallas at least $25\,\unit{\%}$ (3.~quantile) of anomalies are significant on a $3\sigma$ level or more, while at Svanvik the 3.~quantile lies at about $2\sigma$. This means that in 2018 the stations at Esrange and Pallas saw more days with elevated ozone than was monitored at Svanvik.
